---
title: "R Notebook"
output: html_notebook
---


1)reading the dataset
```{r}
hcdsd=read.csv("C:/Users/prabh/Downloads/healthcare-dataset-stroke-data (1).csv",stringsAsFactors = TRUE)
hcdsd
```
as id is unique, let's remove "id" variable
```{r}
hcdsd$id=NULL
```
Let's see the structure and summary of the dataset
```{r}
str(hcdsd)
summary(hcdsd)
```
```{r}
sapply(colnames(hcdsd), function(x) class(hcdsd[[x]]))
which(is.na(hcdsd))
colSums(is.na(hcdsd))
```
ans: There are 5 categorical variables and 7 numeric variables(including "id").
     There are 201 missing values in the variable- "bmi".

```{r}
prop.table(table(hcdsd$stroke))
```

let's convert the stroke variable into factor and explore it by seeing the percentage of each level.
```{r}
hcdsd$stroke=factor(hcdsd$stroke)
levels(hcdsd$stroke)=list(No ="0", Yes="1")
prop.table(table(hcdsd$stroke))

```
By this, we can understand that the stroke(output) variable is imbalanced.
1 if the patient had a stroke - 4%
0 if the patient did not had a stroke - 95%

```{r}
hcdsd$hypertension=factor(hcdsd$hypertension, levels=c("0","1"), labels = c("No", "Yes"))
hcdsd$heart_disease=factor(hcdsd$heart_disease, levels=c("0","1"), labels = c("No", "Yes"))
```

Exploratory data analysis
```{r}
table1<-table(hcdsd$stroke,hcdsd$gender)
mosaicplot(table1, main = "Mosaic plot of stroke VS gender", shade = TRUE)
chisq.test(table1)
table2<-table(hcdsd$stroke,hcdsd$hypertension)
mosaicplot(table2, main = "Mosaic plot of stroke VS hypertension", shade = TRUE)
chisq.test(table2)
table3<-table(hcdsd$stroke,hcdsd$heart_disease)
mosaicplot(table3, main = "Mosaic plot of stroke VS heart_disease", shade = TRUE)
chisq.test(table3)
table4<-table(hcdsd$stroke,hcdsd$ever_married)
mosaicplot(table4, main = "Mosaic plot of stroke VS ever_married", shade = TRUE)
chisq.test(table4)
table5<-table(hcdsd$stroke,hcdsd$work_type)
mosaicplot(table5, main = "Mosaic plot of stroke VS work_type", shade = TRUE)
chisq.test(table5)
table6<-table(hcdsd$stroke,hcdsd$Residence_type)
mosaicplot(table6, main = "Mosaic plot of stroke VS Residence_type", shade = TRUE)
chisq.test(table6)
table7<-table(hcdsd$stroke,hcdsd$smoking_status)
mosaicplot(table7, main = "Mosaic plot of stroke VS smoking_status", shade = TRUE)
chisq.test(table7)
plot(hcdsd$age~hcdsd$stroke)
t.test(hcdsd$age~hcdsd$stroke)
plot(hcdsd$avg_glucose_level~hcdsd$stroke)
t.test(hcdsd$avg_glucose_level~hcdsd$stroke)
plot(hcdsd$bmi~hcdsd$stroke)
t.test(hcdsd$bmi~hcdsd$stroke)
```
ans: variables- hypertension, heart_disease, ever_married, work_type, smoking_status, age, avg_glucose_level, bmi are associated with "stroke" variable

Using caret's createDataPartition method to partition the dataset to 80% training and 20% testing.
```{r}
library(caret)
intrain= createDataPartition(hcdsd$stroke, p=0.8, list= FALSE)
hcdsd_train= hcdsd[intrain,]

hcdsd_test = hcdsd[-intrain,]
```

```{r}
hcdsd_train$bmi[is.na(hcdsd_train$bmi)] <- mean(hcdsd_train$bmi,na.rm=TRUE)
hcdsd_test$bmi[is.na(hcdsd_test$bmi)] <- mean(hcdsd_train$bmi,na.rm=TRUE)
colSums(is.na(hcdsd_train))
colSums(is.na(hcdsd_test))
```

```{r}
str(hcdsd_train)
```
```{r}
str(hcdsd_train$stroke)
```



Lasso Linear Regression model 
```{r}
library(caret)
library(RANN)
library(ROCR)
set.seed(1)
lasso <-train(stroke ~ ., data = hcdsd_train, method = "glmnet",verbose=FALSE, metric ="ROC",
trControl= trainControl("cv", number = 10, classProbs= TRUE, summaryFunction= twoClassSummary, sampling="smote"),
tuneGrid= expand.grid(alpha = 1, lambda = 10^seq(-3, 3, length = 100)))

coef(lasso$finalModel, lasso$bestTune$lambda)

predictions <-predict(lasso,hcdsd_test,na.action = na.pass, type="prob")
lasso_predictions= prediction(predictions$Yes,hcdsd_test$stroke)
lasso_predicted_labels= predict(lasso, hcdsd_test)
confusionMatrix(lasso_predicted_labels, hcdsd_test$stroke, positive="Yes", mode="everything")
performance(lasso_predictions, measure = "auc")@y.values
```


Ridge Linear Regression model
```{r}
set.seed(1)
ridge <-train(stroke ~ ., data = hcdsd_train, method = "glmnet", na.action=na.pass, preProc="knnImpute",verbose=FALSE, metric ="ROC",
trControl= trainControl("cv", number = 10,classProbs= TRUE, summaryFunction= twoClassSummary, sampling="down"),
tuneGrid= expand.grid(alpha = 0, lambda = 10^seq(-3, 3, length = 100)))

predictions1 <-predict(ridge,hcdsd_test,na.action=na.pass, type="prob")
ridge_predictions= prediction(predictions1$Yes,hcdsd_test$stroke)
ridge_predicted_labels= predict(ridge, hcdsd_test)
confusionMatrix(ridge_predicted_labels, hcdsd_test$stroke, positive="Yes", mode="everything")
performance(ridge_predictions, measure = "auc")@y.values
```


Elastic-net Linear Regression model
```{r}
set.seed(1)
enet<-train(stroke ~ ., data = hcdsd_train, method = "glmnet", na.action=na.pass, preProc="knnImpute",verbose=FALSE, metric ="ROC",
trControl= trainControl("cv", number = 10,classProbs= TRUE, summaryFunction= twoClassSummary, sampling="down"),
tuneGrid= expand.grid(alpha =seq(0,1, length=10), lambda = 10^seq(-3, 3, length = 100)))

predictions2 <-predict(enet,hcdsd_test,na.action=na.pass, type="prob")
enet_predictions= prediction(predictions2$Yes,hcdsd_test$stroke)
enet_predicted_labels= predict(enet, hcdsd_test)
confusionMatrix(enet_predicted_labels, hcdsd_test$stroke, positive="Yes", mode="everything")
performance(enet_predictions, measure = "auc")@y.values

```

Random forest model
```{r}
set.seed(1)
library(caret)
m_rf<-train(stroke ~ ., data = hcdsd_train,preProc=c("nzv","knnImpute"),importance=T,method = "rf", verbose=FALSE, metric ="ROC", trControl= trainControl(method = "cv", number = 10,classProbs= TRUE, summaryFunction= twoClassSummary, sampling="down"))
m_rf
predictions3<-predict(m_rf, hcdsd_test, type = "prob")
rf_predictions= prediction(predictions3$Yes,hcdsd_test$stroke)
rf_predicted_labels= predict(m_rf, hcdsd_test)
confusionMatrix(rf_predicted_labels, hcdsd_test$stroke, positive="Yes", mode="everything")
performance(rf_predictions, measure = "auc")@y.values
```
```{r}
varImp(m_rf)
```
```{r}
t=table(hcdsd_train$stroke)
t
w_yes=sum(t)/(t["Yes"]*2)
w_no=sum(t)/(t["No"]*2)
w_yes
w_no
class_weights= ifelse(hcdsd_train$stroke=="Yes", w_yes, w_no)
str(class_weights)
```

Gradient Boosted Tree model
```{r}
library(caret)
library(ROCR)
set.seed(1)
ctrl=trainControl(method="cv", number=10, classProbs= TRUE, summaryFunction= twoClassSummary,sampling="down")
gbm<-train(stroke ~ ., data = hcdsd_train, method = "gbm",preProc="nzv", verbose=FALSE, metric ="ROC", trControl= ctrl)
gbm
gbm_predictions_prob=predict(gbm, hcdsd_test, type="prob")
gbm_predictions= prediction(gbm_predictions_prob$Yes,hcdsd_test$stroke)
performance(gbm_predictions, measure = "auc")@y.values
gbm_predicted_labels= predict(gbm, hcdsd_test)
confusionMatrix(gbm_predicted_labels, hcdsd_test$stroke, positive="Yes", mode="everything")
```

Support vector machine model with method="svmLinear"
```{r}
set.seed(1)
library(caret)
svm<-train(
stroke ~ ., data = hcdsd_train, method = "svmLinear",verbose=FALSE, metric ="ROC",
trControl= trainControl("cv", number = 10, classProbs= TRUE, summaryFunction= twoClassSummary,sampling="down"))
svm
predictions4<-predict(svm,hcdsd_test,type="prob")
svm_predictions= prediction(predictions4$Yes,hcdsd_test$stroke)
svm_predicted_labels= predict(svm, hcdsd_test)
confusionMatrix(svm_predicted_labels, hcdsd_test$stroke, positive="Yes", mode="everything")
performance(svm_predictions, measure = "auc")@y.values
```

Support vector machine model with method="svmRadial"
```{r}
library(caret)
set.seed(1)
svm_rad<-train(
stroke ~ ., data = hcdsd_train, method = "svmRadial",verbose=FALSE, metric ="ROC",
trControl= trainControl("cv", number = 10,classProbs= TRUE, summaryFunction= twoClassSummary,sampling="down"))
svm_rad
predictions4_rad<-predict(svm_rad,hcdsd_test,type="prob")
svm_rad_predictions= prediction(predictions4_rad$Yes,hcdsd_test$stroke)
svm_rad_predicted_labels= predict(svm_rad, hcdsd_test)
confusionMatrix(svm_rad_predicted_labels, hcdsd_test$stroke, positive="Yes", mode="everything")
performance(svm_rad_predictions, measure = "auc")@y.values
```


```{r}
compare=resamples(list(L=lasso,R=ridge,E=enet,RF=m_rf,G=gbm,S=svm,SR=svm_rad))
summary(compare)
```
```{r}
str(hcdsd_train)
str(hcdsd_test)
```
Neural networks require a numeric target variable. In case of a binary classification, the target variable must be coded as 0 and 1. so, converting stroke into numeric.
```{r}
#hcdsd_train$stroke=as.integer(hcdsd_train$stroke)
hcdsd_train$stroke=as.numeric(hcdsd_train$stroke)-1
hcdsd_train$hypertension=as.numeric(hcdsd_train$hypertension)
hcdsd_train$heart_disease=as.numeric(hcdsd_train$heart_disease)

hcdsd_test$stroke=as.numeric(hcdsd_test$stroke)-1
hcdsd_test$hypertension=as.numeric(hcdsd_test$hypertension)
hcdsd_test$heart_disease=as.numeric(hcdsd_test$heart_disease)
```
The categorical variables ever_married, Residence_type have only two levels so we just convert it to a binary vector
```{r}
hcdsd_train$ever_married=ifelse(hcdsd_train$ever_married=="No",0,1)
hcdsd_train$Residence_type=ifelse(hcdsd_train$Residence_type=="Rural",0,1)

hcdsd_test$ever_married=ifelse(hcdsd_test$ever_married=="No",0,1)
hcdsd_test$Residence_type=ifelse(hcdsd_test$Residence_type=="Rural",0,1)
```
The rest of the categorical variables (gender, work_type,smoking_status) don’t have a natural ordering so we one-hot-encode them.
```{r}
library(mltools)
library(data.table)
hcdsd_train= data.frame(one_hot(as.data.table(hcdsd_train),dropUnusedLevels = FALSE))
hcdsd_test= data.frame(one_hot(as.data.table(hcdsd_test),dropUnusedLevels = FALSE))
```
```{r}
str(hcdsd_train)
str(hcdsd_test)
```
```{r}
hcdsd_train
```



```{r}
numeric_cols=c("age","avg_glucose_level","bmi")
col_means_train<-attr(scale(hcdsd_train[,numeric_cols]), "scaled:center")
col_stddevs_train<-attr( scale(hcdsd_train[,numeric_cols]), "scaled:scale")
```
down sampling for neural networks
```{r}
hcdsd_train1= hcdsd_train[hcdsd_train$stroke==0,]
hcdsd_train1
hcdsd_train2= hcdsd_train[hcdsd_train$stroke==1,]
hcdsd_train2
```
```{r}
train_sample<-sample(3889, 200)
hcdsd_train3<-hcdsd_train1[train_sample,]
hcdsd_train3
```
```{r}
hcdsd_train4<-rbind(hcdsd_train3, hcdsd_train2)
hcdsd_train4
```


```{r}
library(caret)
intrain1= createDataPartition(hcdsd_train$stroke, p=0.9, list= FALSE)
hcdsd_train_x= hcdsd_train[intrain1, -20]
hcdsd_train_y= hcdsd_train[intrain1, 20]

hcdsd_val_x= hcdsd_train[-intrain1, -20]
hcdsd_val_y= hcdsd_train[-intrain1, 20]
```
```{r}
numeric_cols=c("age","avg_glucose_level","bmi")
col_means_train1<-attr(scale(hcdsd_train4[,numeric_cols]), "scaled:center")
col_stddevs_train1<-attr( scale(hcdsd_train4[,numeric_cols]), "scaled:scale")
```

```{r}
hcdsd_train_x[numeric_cols]= scale(hcdsd_train_x[numeric_cols])
hcdsd_val_x[numeric_cols]= scale(hcdsd_val_x[numeric_cols], center = col_means_train, scale = col_stddevs_train)
hcdsd_test[numeric_cols]<-scale(hcdsd_test[numeric_cols], center = col_means_train, scale = col_stddevs_train)
```

```{r}
str(hcdsd_train_x)
```

```{r}
library(keras)
model <-keras_model_sequential() %>%
layer_dense(units = 32, activation = "relu",
input_shape= dim(hcdsd_train_x)[2]) %>%
layer_dense(units = 32,activation="relu" )%>%
layer_dense(units = 1, activation="sigmoid")
model %>% compile(
loss = "binary_crossentropy",
optimizer = "adam" , metrics="AUC")
model %>% fit(as.matrix(hcdsd_train_x),hcdsd_train_y,batch_size=50,epochs = 20,verbose=2, validation_data=list(as.matrix(hcdsd_val_x),hcdsd_val_y),class_weight=list("0"=w_no, "1"=w_yes))
```
```{r}
str(hcdsd_train_x)
```

```{r}
str(hcdsd_test)
```


```{r}
predicted_probs= model%>% predict(as.matrix(hcdsd_test[,-20]))
ann_predictions= prediction(predicted_probs,hcdsd_test[,20])
performance(ann_predictions, measure = "auc")@y.values
```

```{r}
predicted.labels= factor(ifelse(predicted_probs>0.5, "1", "0"))
confusionMatrix(predicted.labels, as.factor(hcdsd_test[,20]), mode="everything", positive="1")
```

```{r}
library(tfruns)
 runs=tuning_run("hcdsd_hyperparameters.R",
            flags = list(
              learning_rate = c(0.1, 0.01, 0.001, 0.0001),
              batch_size = c(32,32,1),
              activation_function = c("relu","sigmoid","tanh"),
              units1= c(16,32,64,128),
              units2= c(16,32,64,128)
            ),
            sample = 0.01 
 )
```

```{r}
runs=runs[order(runs$metric_val_auc, decreasing=TRUE),]
runs
```
```{r}
view_run(runs$run_dir[4])
```
```{r}
hcdsd_train_new=rbind(hcdsd_train_x,hcdsd_val_x)
hcdsd_train_newlabels=c(hcdsd_train_y,hcdsd_val_y)
```


```{r}
model <-keras_model_sequential() %>%
layer_dense(units = 128, activation = "relu",
input_shape= dim(hcdsd_train_new)[2]) %>%
layer_dense(units = 128,activation="relu" )%>%
layer_dense(units = 1, activation="sigmoid")
model %>% compile(
loss = "binary_crossentropy",
optimizer = optimizer_adam(learning_rate=0.0001) , metrics="AUC")
model %>% fit(as.matrix(hcdsd_train_new),hcdsd_train_newlabels,batch_size=32,epochs = 20,verbose=2, validation_data=list(as.matrix(hcdsd_val_x),hcdsd_val_y),class_weight=list("0"=w_no, "1"=w_yes))
```
```{r}
summary(model)
```
```{r}
str(hcdsd_train_new)
str(hcdsd_test)
```


```{r}
library(caret)
predicted_probs= model%>% predict(as.matrix(hcdsd_test[,-20]))
predicted.labels= factor(ifelse(predicted_probs>0.5, "1", "0"))
confusionMatrix(predicted.labels, as.factor(hcdsd_test[,20]), mode="everything", positive="1")
```
Plots of various models used in this project

```{r}
plot(lasso)
```
```{r}
plot(ridge)
```

```{r}
plot(enet)
```

```{r}
plot(m_rf)
```
```{r}
plot(gbm)
```
```{r}
plot(svm_rad)
```



